Seminarium TI 3700\\ Data Analysis on Twitter: An Overview (working title)

### 2014-03-10 authorinstr_2.tex
Kleine toevoegingen
API revised

When Twitter was conceived, the developers decided that to adopt an API (Application Specific Interface) for third party developers. This has proven to be very
beneficial to both Twitter's popularity and alot of its users, including researchers.

An overview of the API structure and some useful fields:
\begin{enumerate}
\item \textbf{Tweets} The short messages sent by users
	\subitem \textbf{Coordinates} Retrieves the longitude and latitude position from where a specified tweet has been sent
	\subitem \textbf{Entities} Retrieves the entities (url, hashtag, etc.) that are embedded into the tweet
	\subitem \textbf{Lang} Retrieves the language in which a tweet has been sent
	\subitem \textbf{User} Retrieves the user that sent a tweet
\item \textbf{Users} The accounts that send tweets
	\subitem \textbf{Description} Retrieves the description that the user defined for his account
	\subitem \textbf{Location} Retrieves the location that the user defined for his account
	\subitem \textbf{Status} Retrieves, if possible, the user's latest tweet
\item \textbf{Entities} Images, hashtags and other entities that tweets can hold
	\subitem \textbf{Hashtags} Retrieves the hashtags that have been parset out of a tweet
	\subitem \textbf{Media} Retrieves the media that is associated with a tweet
	\subitem \textbf{urls} Retrieves the urls embedded in a tweet
	\subitem \textbf{user\_mentions} Retrieves the users that have been mentioned in a tweet
\item \textbf{Places} Predefined places on Twitter that users can associate with
	\subitem \textbf{Country} Retrieves the country associated with a place
	\subitem \textbf{Full\_name} Retrieves the full name of a place, for example Kanses ``City, Kansas''
	\subitem \textbf{Place\_type} Retrieves the type of a place, for example ``city''
\end{enumerate}

When the above fields are combined, it is possible to create sets of Tweets that fulfill a certain specification. These sets can be used to extract information.
No matter how specific you can define your search, sometimes you need a random sample of tweets to analyse. For this purpose, Twitter has a stream that
third party developers can tap into to retrieve random tweets. Unfortunately, the Twitter stream can only retrieve tweets that are less than a few days old.

@@

\section{Single tweet analysis}
In this section we will discuss the different types of analysis that can be done on one single tweet. This usually is also the basis for a analysis of a larger amount of tweets or even on a stream of tweets. 
\subsection{Data enrichment}
At the very basis of the analysis of a tweet is data enrichment. Because the length of a tweet is restricted to 140 characters, the users of Twitter use language in creative ways. This way of communicating makes it sometimes, even for humans, hard to understand what a tweet is about. To get a better understanding of the meaning of a tweet, a context is needed. For humans this is done by linking the concepts from the tweet to earlier acquired knowledge. The same thing can be done by computers. Two stages can be defined in this process, first the extraction of the concepts from the tweet, after that a context can be created by linking the tweet to other sources.
\subsubsection{Concept extraction}
In this case by concepts we mean named entities such as names of locations, people and organizations. To recognize these entities multiple learning algorithms or even combinations of learning algorithms have been proposed. The most common is statistical learning. Think of algorithms like Maximum Entropy Models, Hidden Markov Models or neural networks. Other proposed algorithms are AdaBoost.MH, employed memory-based learning, transformation-based learning, support vector machines and conditional random fields. All of them in decreasing frequency of use.\cite{EntityRecognition} Florian et al. have shown that combining multiple classifiers can increase the performance, compared to the best performing classifier, by 17-21 \%.\cite{ClassifierCombination} After the entities have been found, labels are assigned to distinguish between a person, an organization or location. Then Abel et al. propose to use the RDF.\cite{AdaptiveSearch} By doing so sources from the Semantic Web (e.g. DBpedia \footnote{http://dbpedia.org/}) can now be leveraged to enrich the data. Another advantage of using RDF is that RDF is a Semantic Web standard recommended by the W3C\footnote{http://www.w3.org/RDF/}. Because W3C recommends to use this model, most likely a lot of other sources also use this model. This will give a much bigger knowledge base to enrich the data from.
\subsubsection{Context creation}
The problem of linking tweets to another source of information poses the biggest issue in data enrichment. The tweets themselves do not provide much information. Fortunately, Kwak et al. have shown that the topic of over 85\% of the of the tweets is related to news \cite{newsmedia}. This knowledge can be leveraged to create a context for most of the tweets.  This is what Abel et al. proposed to do\cite{enrichmentForProfiling}. They linked tweets to mainstream news networks like CNN, BBC and New York Times.

@@

%TODO misschien wat meer info over het semantic web toevoegen.
After this process \cite{eventalign} uses multiple techniques to align a tweet with the respective event. The first technique used by \cite{eventalign} is Feature Extraction. Three featuresets are used:
\begin{itemize}
%TODO uitwerken
  \item Immediate Resource Leaves: Use only the URI which are most indicative of an event. This feature only uses literals and resources within the URI of an event. 
  \item 1-Step Resource Leaves: Use all literals and resources within 1 step from the URI of an event. 
  \item DBPedia Concepts: Also uses all the concepts from DBPedia 
\end{itemize}
%TODO afmaken

\subsubsection{Usecase: Earthquake Detection}
In \cite{earthq} a usecase for Event Identification is presented in the form of a earthquake detection system. \cite{earthq} considers twitter users as sensors. If enough sensors are activated (Tweets sent by a user are classified as a tweet about an earthquake), it is likely that an earthquake is occurring.

>>>>

When Twitter was conceived, the creators chose to adopt an API (Application Specific Interface) for third party developers. This has proven to be very
beneficial to both Twitter's popularity and alot of its users, including researchers.

An overview of the API structure and some useful fields:
\begin{enumerate}
\item \textbf{Tweets} The short messages consisting of 140 characters sent by users
	\subitem \textbf{Coordinates} Returns the longitude and latitude position from where a specified tweet has been sent
	\subitem \textbf{Entities} Returns the entities (url, hashtag, etc.) that are embedded into the tweet
	\subitem \textbf{Lang} Returns the language in which a tweet has been sent
	\subitem \textbf{User} Returns the user that sent a tweet
\item \textbf{Users} The users that send the tweets
	\subitem \textbf{Description} Returns the description that the user defined for his account
	\subitem \textbf{Location} Returns the location that the user defined for his account
	\subitem \textbf{Status} Returns, if possible, the user's latest tweet
\item \textbf{Entities} Images, hashtags and other entities that tweets can hold
	\subitem \textbf{Hashtags} Returns the hashtags that have been parset out of a tweet
	\subitem \textbf{Media} Returns the media that is associated with a tweet
	\subitem \textbf{urls} Returns the urls embedded in a tweet
	\subitem \textbf{user\_mentions} Returns the users that have been mentioned in a tweet
\item \textbf{Places} Predefined places on Twitter that users can associate with
	\subitem \textbf{Country} Returns the country associated with a place
	\subitem \textbf{Full\_name} Returns the full name of a place, for example Kanses ``City, Kansas''
	\subitem \textbf{Place\_type} Returns the type of a place, for example ``city''
\end{enumerate}

When the above fields are combined, it is possible to create datasets of Tweets that fulfill a certain specification. 
These sets can then be used to extract specific information. However, defining your own dataset may introduce a certain bias, 
and sometimes you need a random sample of tweets to analyse. For this purpose, Twitter has a streaming function that
third party developers can tap into to retrieve random tweets. Unfortunately, the Twitter stream can only retrieve tweets that are less than a few days old.

@@

\section{Single tweet analysis}
In this section we will discuss the different types of analysis that can be done on one single tweet. This usually is also the basis for a analysis of a larger amount of tweets or even on a stream of tweets. 
\subsection{Data enrichment}
At the very basis of the analysis of a tweet is data enrichment. Because the length of a tweet is restricted to 140 characters, the users of Twitter use language in creative ways. This way of communicating makes it sometimes, even for humans, hard to understand what a tweet is about. To get a better understanding of the meaning of a tweet, a context is needed. For humans this is done by linking the concepts from the tweet to earlier acquired knowledge. The same thing can be done by computers. Two stages can be defined in this process, first the extraction of the concepts from the tweet, after that a context can be created by linking the tweet to other sources.
\subsubsection{Concept extraction}
In this case by concepts we mean named entities such as names of locations, people and organizations. To recognize these entities multiple learning algorithms or even combinations of learning algorithms have been proposed. The most common is statistical learning. Think of algorithms like Maximum Entropy Models, Hidden Markov Models or neural networks. Other proposed algorithms are AdaBoost.MH, employed memory-based learning, transformation-based learning, support vector machines and conditional random fields. All of them in decreasing frequency of use.\cite{EntityRecognition} Florian et al. have shown that combining multiple classifiers can increase the performance, compared to the best performing classifier, by 17-21 \%.\cite{ClassifierCombination} After the entities have been found, labels are assigned to distinguish between a person, an organization or location. Then Abel et al. propose to use the RDF.\cite{AdaptiveSearch} By doing so sources from the Semantic Web (e.g. DBpedia \footnote{http://dbpedia.
org/}) can now be leveraged to enrich the data. Another advantage of using RDF is that RDF is a Semantic Web standard recommended by the W3C\footnote{http://www.w3.org/RDF/}. Because W3C recommends to use this model, most likely a lot of other sources also use this model. This will give a much bigger knowledge base to enrich the data from.
\subsubsection{Context creation}
The problem of linking tweets to another source of information poses the biggest issue in data enrichment. The tweets themselves do not provide much information. Fortunately, Kwak et al. have shown that the topic of over 85\% of the of the tweets is related to news \cite{newsmedia}. This knowledge can be leveraged to create a context for most of the tweets.  This is what Abel et al. proposed to do\cite{enrichmentForProfiling}. They linked tweets to mainstream news networks like CNN, BBC and New York Times.

@@

%TODO misschien wat meer info over het semantic web toevoegen.
After this process \cite{eventalign} uses multiple techniques to align a tweet with the respective event. The first technique used by \cite{eventalign} is Feature Extraction. Three featuresets are used:
\begin{itemize}
%TODO uitwerken
  \item Immediate Resource Leaves: Use only the URI which are most indicative of an event. This feature only uses literals and resources within the URI of an event. 
  \item 1-Step Resource Leaves: Use all literals and resources within 1 step from the URI of an event. 
  \item DBPedia Concepts: Also uses all the concepts from DBPedia 
\end{itemize}
After this a Feature Vector is composed, for tweets all stop words are removed from the bag of words a tweet is represented as. Furthermore all stopwords are removed and everything is normalized to lowercase. \cite{eventalign} Then a labeling function is used to label the event. 
\subsubsection{Usecase: Earthquake Detection}
In \cite{earthq} a usecase for Event Identification is presented in the form of a earthquake detection system. \cite{earthq} considers twitter users as sensors. If enough sensors are activated (Tweets sent by a user are classified as a tweet about an earthquake), it is likely that an earthquake is occurring.
